---
title: "PA_template"
author: "Paz Prieto Martín"
date: "17 de julio de 2014"
output: html_document
---

Loading and preprocessing the data
---
# Load the data
```{r}
activity <- read.csv("~/Desktop/COURSERA R/Reproducible Research/activity.csv", stringsAsFactors=TRUE)
```
---
What is mean total number of steps taken per day?
---
# Histogram of the total number of steps taken each day:
```{r}
activity$steps<-as.integer(activity$steps)
hist(with(activity, tapply(steps, list(date), sum)),breaks=61)
```
# Mean of steps taken each day:
```{r}
tapply(activity$steps, activity$date,mean)
activityNoNA<-activity[complete.cases(activity),]
meanNA<-tapply(activityNoNA$steps, activityNoNA$date,mean)
totalmeanwithNA<-mean(meanNA[complete.cases(meanNA)])
totalmeanwithNA

```
# Median number of steps taken each day:
```{r}
medianNA<-tapply(activityNoNA$steps, activityNoNA$date,median)
totalmedianwithNA<-median(medianNA[complete.cases(medianNA)])
totalmedianwithNA
```
What is the average daily activity pattern?
---
# time series plot of the average number of steps taken (averaged across all days) versus the 5-minute intervals (5-minute interval (x-axis),average number of steps taken, averaged across all days (y-axis)):
```{r}
activityNoNA$interval<-as.factor(activityNoNA$interval)
activityNoNA$averageacrossdays<-tapply(activityNoNA$steps, activityNoNA$interval,mean)
plot(activityNoNA$interval,activityNoNA$averageacrossdays, type="l", xlab="5-minute interval", ylab="number of steps averaged across all days", col="green", lwd=2)
```
# 5-minute interval that, on average, contains the maximum number of steps:
```{r}
activityNoNA$interval[which.max(activityNoNA$averageacrossdays)]
```
Imputing missing values
----
# Total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
nrow(activity)- nrow(activityNoNA)
```
# Strategy for imputing missing data:
# Replace each NA with the average number of steps taken across all days for the specific interval. 
# actReplNA is a new dataset thaat is equal to the original data but with the missing data filled in.
```{r}
actReplNA<-activity
actReplNA$interval<- as.factor(actReplNA$interval)
actReplNA$steps<-ifelse(is.na(actReplNA$steps),activityNoNA$averageacrossdays,actReplNA$steps)
summary(is.na(actReplNA$steps))
````
# Histogram of the total number of steps taken each day after missing values were imputed:
```{r}
hist(with(actReplNA, tapply(steps, list(date), sum)),breaks=61)
````
# Calculate and report the mean total number of steps taken per day:
```{r}
meanReplNA<-tapply(actReplNA$steps, actReplNA$date,mean)
totalmeanReplNA<-mean(meanReplNA)
totalmeanReplNA

```
# Calculate and report the median total number of steps taken per day:
```{r}
medianReplNA<-tapply(actReplNA$steps, actReplNA$date,median)
totalmedianReplNA<-median(medianReplNA)
totalmedianReplNA
```

#Do these values differ from the estimates from the first part of the assignment? 
```{r}
DifferMean<-meanReplNA-meanNA
DifferMean
DifferMedian<-medianReplNA-medianNA
DifferMedian
```
No differences were observed by replacing NA's in mean or median. But it was possible to estimate the total daily number of steps for the dates without steps data:
```{r}
sumwithNA<-tapply(activityNoNA$steps, activityNoNA$date,sum)
sumwithNA
sumReplNA<-tapply(actReplNA$steps, actReplNA$date,sum)
sumReplNA
```
---
Are there differences in activity patterns between weekdays and weekends?
---
# Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends:
```{r}
actReplNA$daytype<-weekdays(as.Date(actReplNA$date))
actReplNA$daytype<-as.factor(actReplNA$daytype)
actReplNA$day<-factor(c("Weekday","Weekend"))
for (i in 1:17568)
        {
        daytype <- actReplNA$daytype[i]
        
        if ((daytype =="sábado") | (daytype=="domingo"))
                {
                actReplNA$day[i]<-"Weekend"
                }
        else 
                {
                actReplNA$day[i]<-"Weekday"
                 }
      }  
dataW<-split(actReplNA, actReplNA$day)
Weekday<-(dataW[[1]])
Weekend<-(dataW[[2]])
Weekday$averageacross<-tapply(Weekday$steps, Weekday$interval,mean)

Weekend$averageacross<-tapply(Weekend$steps, Weekend$interval,mean)
Data<-rbind(Weekday,Weekend)

library(lattice)
xyplot(averageacross~interval | day, data=Data, layout=c(1,2), type="l", xlab="5-minute interval", ylab="number of steps")
````
